From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-02 Pi+/paper/equivalence.tex
Add some text in section 6

In this section, we consider the additive fragment $\PiPlusLang$ of $\PiLang$. This is the language that we interpret to
$\UFin$ and back, using the tools developed in the previous sections. Further, we go through an intermediate step of the
language $\PiHatLang$, which is a simplified variant of $\PiPlusLang$ that uses adjacent transpositions for combinators,
but still preserving all the required structure.

@@

Next, we show how to translate $\PiPlusLang$ programs to $\PiHatLang$ and back. $\PiHatLang$ is a simplified variant of
$\PiPlusLang$, with (unary) natural numbers for types, constructed by $\zero$ and $\suc$, 1-combinators generated by
adjacent transpositions, and an appropriate set of 2-combinators. We give the syntax, again omitting 2-combinators,
in~\cref{fig:pihat}. \vc{Something about PROPs.}

@@

As described, $\PiHatLang$ doesn't have a tensor product, but we can build it simply by adding up natural numbers.
However, we need to verify that this indeed equips $\PiHatCat$ with a symmetric monoidal strucuture.

@@

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. The
non-trivial part is coherently generating a $\PiPlusLang$ combinator from $\PiHatLang$ combinators, using adjacent swaps
to generate big swaps.

@@

Types in $\PiHatLang$ are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The
1-combinators in $\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, or permutations, in
each connected component. Finally, 2-combinators are mapped to 2-paths between loops in $\UFin$.

@@

0-cells in $\UFin$ are mapped to their cardinalities in $\PiHatLang$, 1-loops are decoded to words in $\Sn$ to generate
a sequence of adjacent transpositions, producing a 1-combinator in $\PiHatLang$. 2-paths are quoted back to
2-combinators in $\PiHatLang$.

@@

$\UFin$ as a 1-groupoid, is a (locally strict) weak $(2,0)$-category.

@@

\subsection{$\PiLang$ circuit normalization}

To normalise a $\PiLang$ circuit, we translate it to $\PiHatLang$, compute a permutation, and quote it back to
$\PiPlusLang$. Note that the normalisation happens in going from $\PiHatLang$ to $\UFin$ to $\PiHatLang$.

>>>>

In this section, we consider the additive fragment $\PiPlusLang$ of $\PiLang$. This is the language that we interpret to
$\UFin$ and back, using the tools developed in the previous sections. Further, we go through an intermediate step of the
language $\PiHatLang$, which is a simplified variant of $\PiPlusLang$ that uses adjacent transpositions for combinators,
while preserving all the required structure.

@@

Next, we show how to translate $\PiPlusLang$ programs to $\PiHatLang$ and back. $\PiHatLang$ is a simplified variant of
$\PiPlusLang$, with (unary) natural numbers for 0-cells, 1-combinators generated by adjacent transpositions, and an
appropriate set of 2-combinators. We give the syntax, again omitting 2-combinators, in~\cref{fig:pihat}.

@@

As described, $\PiHatLang$ doesn't have a tensor product, but we can build it simply by adding up natural numbers, and,
we need to verify that this indeed equips $\PiHatCat$ with a symmetric monoidal strucuture. Since each object is a
natural number generated by $\zero$ and $\suc$, this makes $\PiHatCat$ similar to a \emph{PROP}, that is, a products and
permutations category.

@@

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. The
non-trivial part is coherently generating a $\PiPlusLang$ combinator from $\PiHatLang$ combinators, using adjacent swaps
to generate the braiding.

@@

Types in $\PiHatLang$ are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The
1-combinators in $\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, in each connected
component. In $\PiHatLang$, the 1-combinators are generated by adjacent transpositions, so these can be easily mapped to
words in $\Sn$ and then to automorphisms using~\cref{prop:sn-lehmer-fin-equiv}. Finally, 2-combinators are mapped to
2-paths between loops in $\UFin$.

@@

0-cells in $\UFin$ are mapped to their cardinalities in $\PiHatLang$, 1-loops are decoded to words in $\Sn$ to generate
a sequence of adjacent transpositions, producing a 1-combinator in $\PiHatLang$. Finally, 2-paths are quoted back to
2-combinators in $\PiHatLang$.

@@

$\UFin$ as a 1-groupoid, is a (locally strict) weak $(2,0)$-category, and we use this to state our final equivalence.

@@

\subsection{$\PiLang$ circuit normalisation}

To normalise a $\PiLang$ circuit, we translate it to $\PiHatLang$, compute a permutation, and quote it back to
$\PiPlusLang$. Note that the normalisation happens in going from $\PiHatLang$ to $\UFin$ and back to $\PiHatLang$.
